% Chapter 3: Literature Review
% Psychological Manipulation in Marketing Discourse
% Target: 6,500 words

\chapter{Literature Review}
\label{ch:literature}

\initial{T}he systematic investigation of psychological manipulation in marketing discourse builds upon decades of interdisciplinary scholarship spanning marketing studies, consumer psychology, discourse analysis, and digital media studies. While extensive research has examined individual components of marketing influence—from persuasion psychology to digital tracking—the integration of these insights into a comprehensive framework for understanding contemporary manipulation remains incomplete. This literature review synthesizes existing scholarship to establish the theoretical foundation upon which this thesis advances our understanding of how marketing discourse systematically exploits consumer vulnerabilities in the digital age.

\section{Introduction: Mapping the Scholarly Landscape}
\label{sec:lit_intro}

The investigation of psychological manipulation in marketing discourse sits at the intersection of multiple scholarly traditions, each contributing unique insights while leaving certain aspects underexplored. This literature review synthesizes research from marketing studies, consumer psychology, discourse analysis, digital media studies, and business ethics to establish the knowledge foundation upon which this thesis builds. The review reveals both the substantial scholarship addressing components of marketing manipulation and the significant gaps that this research addresses.

Three major strands of literature inform this investigation. First, marketing and consumer psychology research has extensively documented how promotional messages influence behavior, though often from an industry perspective prioritizing effectiveness over ethics. Second, discourse analysis scholarship has examined power relations in commercial communication, though with limited attention to psychological mechanisms and digital contexts. Third, emerging literature on digital manipulation addresses technological affordances for influence, though often without the linguistic and semiotic depth necessary for comprehensive understanding.

The review follows a thematic rather than chronological organization, reflecting the interdisciplinary nature of marketing manipulation. After examining foundational work on persuasion and influence, it explores sector-specific studies revealing unique manipulation patterns in fashion, fitness, and skincare markets. The review then addresses digital transformation of marketing practices before identifying critical gaps that this thesis addresses. Throughout, the emphasis remains on synthesizing insights rather than merely cataloging studies, building toward an integrated understanding that transcends disciplinary boundaries.

\section{Foundations: Persuasion, Influence, and Manipulation}
\label{sec:foundations}

\subsection{Classical Persuasion Theory and Marketing Applications}

The study of persuasion in marketing contexts has deep roots extending to Aristotle's rhetoric, but modern scientific investigation began with Carl Hovland's Yale Communication Research Program in the 1950s. Hovland's systematic examination of source credibility, message characteristics, and audience factors established the experimental paradigm that continues to dominate persuasion research. His finding that high-credibility sources produce greater attitude change provided theoretical foundation for authority-based marketing strategies, explaining why brands invest heavily in expert endorsements and scientific credentials.

The Elaboration Likelihood Model (ELM), developed by Richard Petty and John Cacioppo (1986), revolutionized understanding of how persuasion operates through dual processing routes. The central route involves careful evaluation of message arguments, while the peripheral route relies on cognitive shortcuts and emotional responses. Their research demonstrates that peripheral processing dominates under conditions of low motivation, limited ability, or time pressure—conditions that modern marketing deliberately creates. Our empirical finding of temporal pressure tactics in 142 instances across brands confirms systematic exploitation of peripheral processing vulnerabilities.

McGuire's (1989) information processing model identified six stages of persuasion: exposure, attention, comprehension, yielding, retention, and action. Marketing manipulation operates by optimizing each stage while obscuring the persuasive intent. Digital tracking ensures exposure through retargeting, visual design captures attention, simplified messages aid comprehension, emotional appeals facilitate yielding, repetition ensures retention, and convenience features enable action. This stage-based understanding reveals manipulation as systematic process rather than single technique.

\subsection{Cialdini's Influence Principles: From Description to Exploitation}

Robert Cialdini's "Influence: The Psychology of Persuasion" (1984, revised 2021) transformed both academic understanding and practical application of influence principles. His identification of six universal principles—reciprocity, commitment/consistency, social proof, authority, liking, and scarcity—provided actionable framework that marketers rapidly adopted. While Cialdini intended to help consumers recognize and resist manipulation, his work inadvertently became manual for sophisticated exploitation.

Subsequent research has refined understanding of each principle's operation. Goldstein et al. (2008) demonstrated that social proof effectiveness depends on similarity between observer and observed—explaining why brands invest in micro-influencers who resemble target consumers rather than distant celebrities. Griskevicius et al. (2009) showed that scarcity effects intensify under social competition conditions, explaining luxury brands' emphasis on exclusive access. Lynn (1991) revealed that scarcity increases desirability independent of quality changes, confirming that manipulation operates through psychological rather than rational mechanisms.

The digital transformation has amplified these principles' power. A/B testing enables optimization of authority cues, algorithmic recommendation systems weaponize social proof, and dynamic pricing exploits scarcity perception. Guadagno et al. (2013) found that online contexts actually intensify some influence principles—particularly social proof and authority—because uncertainty increases reliance on cognitive shortcuts. Our analysis confirms this intensification: social proof appears 127 times and authority appeals 175 times across digital marketing texts.

\subsection{The Manipulation Distinction: When Influence Becomes Exploitation}

The boundary between legitimate persuasion and unethical manipulation remains contested in literature. Beauchamp (1984) proposed that manipulation involves deception, pressure, or playing upon emotional weakness to subvert rational decision-making. This definition, while useful, struggles with borderline cases where emotional appeals accompany factual information. Our framework addresses this by focusing on exploitation of vulnerabilities against consumer interests.

Susser, Roessler, and Nissenbaum (2019) provide crucial contemporary framework for understanding digital manipulation. They identify three components: hidden influence (operating outside awareness), exploitation of cognitive biases, and subversion of authentic choice. Their work reveals how digital technologies enable manipulation at unprecedented scale through behavioral tracking, predictive modeling, and personalized targeting. The 1,364 manipulation instances identified in our corpus suggest this digital manipulation has become normalized across sectors.

Mills (1958) distinguished between manipulation and coercion, noting that manipulation preserves illusion of choice while constraining options. This insight proves particularly relevant to digital marketing where choice architectures create false agency. Consumers feel they're making free choices while algorithms curate options, defaults steer decisions, and dark patterns obstruct alternatives. The sophisticated choice manipulation we observe—particularly in subscription models and purchase processes—confirms Mills' theoretical predictions.

\section{Marketing Psychology and Consumer Behavior}
\label{sec:marketing_psych}

\subsection{Emotion in Advertising Effectiveness}

The role of emotion in advertising has attracted extensive research, though much focuses on effectiveness rather than ethics. Holbrook and Batra (1987) identified three emotional dimensions in advertising response: pleasure, arousal, and domination. Their framework helps explain why fear appeals prove universally effective—they create high arousal and low pleasure, motivating action to resolve discomfort. Our finding of 248 fear appeals across sectors aligns with their prediction that negative emotions drive behavior more reliably than positive ones.

Bagozzi, Gopinath, and Nyer (1999) developed comprehensive framework linking emotions to consumer behavior through appraisal theory. They demonstrate that specific emotions trigger predictable behavioral responses: fear motivates protection seeking, envy drives acquisition, and pride encourages display. Marketing manipulation exploits these predictable pathways—skincare brands trigger fear of aging to motivate product purchase, fashion brands elicit envy to drive conspicuous consumption, fitness brands cultivate pride to encourage continued engagement.

Recent neuroscience research provides biological understanding of emotional manipulation. Plassmann et al. (2012) used fMRI scanning to reveal how marketing messages activate reward circuits before conscious evaluation occurs. This pre-conscious processing explains why awareness of manipulation doesn't prevent its effectiveness—emotional responses occur before rational assessment. The multimodal manipulation we observe, particularly combining threatening imagery with reassuring text, exploits this temporal gap between emotional and rational processing.

\subsection{Body Image, Identity, and Consumption}

The relationship between marketing and body image has generated substantial literature, particularly regarding harmful effects on self-perception and mental health. Thompson and Heinberg (1999) demonstrated that exposure to idealized media images increases body dissatisfaction and eating disorder risk. Their work reveals how marketing doesn't merely reflect beauty standards but actively constructs them to create dissatisfaction that products promise to resolve.

Dittmar (2008) developed the Consumer Culture Impact Model showing how material goods become identity markers. Marketing manipulates by positioning products as identity solutions—"be yourself" paradoxically means purchasing mass-produced items. This identity manipulation appears throughout our corpus: fashion brands promise authenticity through consumption, fitness brands equate product use with personal transformation, skincare brands link appearance to self-worth.

Recent research on social media marketing reveals intensified identity manipulation. Perloff (2014) documents how Instagram and TikTok create continuous appearance comparison opportunities, amplifying body dissatisfaction. Influencer marketing exploits parasocial relationships where followers adopt influencers' consumption patterns as identity templates. Our analysis of Gymshark's marketing reveals sophisticated identity manipulation: creating "athlete" identity category that requires product purchase for membership.

\subsection{Neuromarketing and Unconscious Influence}

The emergence of neuromarketing—applying neuroscience to marketing—has revealed unconscious dimensions of consumer influence. Morin (2011) reviews evidence that 95\\% of purchase decisions occur below conscious awareness, suggesting that rational persuasion models fundamentally misunderstand consumer behavior. This finding supports our observation that manipulation often operates through unconscious mechanisms—visual cues, emotional priming, and embodied responses.

Plassmann, Rams\u00f8y, and Milosavljevic (2012) demonstrate that neural responses predict purchase behavior better than self-reported preferences. Their research reveals the "neural focus group"—using brain scanning to optimize marketing messages for maximum unconscious impact. While our research lacks neurological data, the systematic deployment of visual and emotional triggers across brands suggests widespread application of neuromarketing insights.

The ethical implications of neuromarketing remain debated. Wilson et al. (2008) argue that brain-based manipulation threatens consumer autonomy more fundamentally than traditional persuasion because it bypasses conscious defenses. Murphy et al. (2008) propose ethical guidelines including transparency, consumer consent, and vulnerability protection. The complete absence of neuromarketing disclosure in our corpus suggests these ethical guidelines remain aspirational rather than practiced.

\section{Sector-Specific Literature}
\label{sec:sector_lit}

\subsection{Luxury Fashion: Constructing Desire and Distinction}

The luxury fashion literature reveals sophisticated understanding of how brands construct and maintain exclusivity while pursuing mass market growth. Kapferer and Bastien (2014) argue that luxury brands face fundamental paradox: maintaining exclusivity while expanding accessibility. Their analysis explains our finding that fear dominates luxury marketing—exclusivity threats prove more motivating than inclusion promises because they activate loss aversion regarding social status.

Dion and Borraz (2016) examine how luxury retail spaces function as "sacred" environments that transform consumption into quasi-religious experience. Their ethnographic research reveals spatial manipulation techniques: imposing architecture that diminishes consumers, reverential product presentation that suggests precious artifacts, and controlled access that materializes exclusivity. These physical techniques translate to digital spaces through virtual boutiques, limited online availability, and membership-only access.

The role of heritage narratives in luxury manipulation has attracted scholarly attention. Balmer (2013) documents how brands construct fictional histories that create value through temporal distance. "Traditional craftsmanship" and "timeless elegance" claims rarely reflect actual production methods—most luxury goods are mass-produced using modern techniques. Our analysis confirms systematic heritage manipulation: fashion brands reference founding dates, archival designs, and historical associations to justify premium pricing.

Recent research on luxury consumption in digital contexts reveals new manipulation strategies. Kim and Ko (2012) identify five dimensions of luxury social media marketing: entertainment, interaction, trendiness, customization, and word-of-mouth. Each dimension offers manipulation opportunities: entertainment disguises commercial intent, interaction creates parasocial relationships, trendiness exploits FOMO, customization provides false uniqueness, and word-of-mouth generates authentic-seeming endorsement.

\subsection{Fitness Industry: Bodies, Transformation, and Discipline}

The fitness industry literature examines how exercise becomes commercialized through product consumption. Sassatelli (2010) analyzes fitness centers as sites of body discipline where commercial and moral imperatives merge. Her ethnographic work reveals how fitness marketing constructs exercise as moral obligation—laziness becomes character flaw, physical appearance reflects inner worth, and product consumption demonstrates virtue.

Millington (2018) traces the "second fitness boom" enabled by wearable technology and social media. His analysis reveals how digital fitness platforms intensify surveillance and comparison, transforming exercise from personal practice to public performance. The gamification strategies we observe—leaderboards, achievement badges, streak counters—exemplify this transformation. Private bodily experience becomes data-generating activity that enables both self-monitoring and social judgment.

The construction of "transformation" narratives in fitness marketing has received critical attention. Sender and Sullivan (2008) analyze how before/after imagery creates impossible temporal compression—months of gradual change appear instantaneous. Their semiotic analysis reveals visual manipulation techniques: different lighting, posture, and clothing that exaggerate apparent change. Our finding of 91 aspiration appeals in fitness marketing confirms the centrality of transformation mythology.

Recent research on fitness influencers reveals parasocial manipulation mechanisms. Raggatt et al. (2018) document how followers develop emotional attachments to fitness influencers that override critical evaluation. Influencers' apparent authenticity—sharing struggles alongside successes—creates trust that facilitates product promotion. The boundary between inspiration and manipulation blurs when commercial relationships remain undisclosed or when results prove unattainable for average consumers.

\subsection{Skincare and Cosmetics: Science, Beauty, and Aging}

The skincare literature examines how scientific discourse legitimizes beauty consumption. Ringrow (2016) analyzes how "cosmeceutical" category blurs boundaries between cosmetics and pharmaceuticals, allowing beauty products to claim medical benefits without medical regulation. Her discourse analysis reveals systematic appropriation of scientific language—our finding of 125 scientific mimicry instances confirms this pattern continues intensifying.

The construction of aging as pathology requiring intervention has attracted scholarly critique. Calasanti (2007) examines how anti-aging marketing creates "age anxiety" that transforms natural processes into problems requiring solution. The medicalization of appearance—wrinkles become "damage," variation becomes "disorder"—justifies continuous consumption. Our analysis reveals sophisticated problem amplification: normal skin characteristics are photographed under harsh lighting and extreme magnification to appear problematic.

Research on beauty advertising's psychological effects reveals harmful consequences. Richins (1991) demonstrated that exposure to idealized beauty images reduces satisfaction with own appearance. This dissatisfaction proves functional for marketers—contentment doesn't motivate purchase. The progression from problem identification to solution provision that we observe across skincare brands exemplifies this manufactured dissatisfaction.

The digital transformation of beauty marketing through apps and AI has created new manipulation possibilities. Ribeiro et al. (2021) examine how beauty apps use augmented reality to show "improved" versions of users' faces, creating dissatisfaction with unfiltered appearance. These technologies don't merely document current appearance but project "potential" appearance contingent on product purchase. The personalized manipulation possible through facial analysis and AI recommendation exceeds traditional marketing's generic appeals.

\section{Digital Marketing Transformation and Algorithmic Manipulation}
\label{sec:digital_lit}

\subsection{From Mass Marketing to Micro-Targeting}

The shift from mass marketing to personalized micro-targeting represents a fundamental transformation in manipulation capabilities. Zuboff (2019) conceptualizes this as "surveillance capitalism" where behavioral data extraction enables unprecedented prediction and modification of consumer behavior. Her analysis reveals how digital platforms transform human experience into behavioral data, which algorithms process to predict and influence future behavior. This "behavioral futures market" trades in human predictability, with marketing manipulation as primary application.

Turow (2011) documents the evolution of consumer tracking from simple cookies to sophisticated cross-device fingerprinting. His research reveals how seemingly innocuous data points—mouse movements, typing patterns, browsing sequences—combine to create detailed psychological profiles. These profiles enable what he terms "digital discrimination": different consumers receive different prices, products, and persuasive messages based on algorithmic assessment of vulnerability and value. Our finding that brands deploy multiple simultaneous strategies suggests algorithmic optimization of manipulation combinations.

The personalization paradox emerges in recent literature. Aguirre et al. (2015) demonstrate that while personalized advertising can increase relevance, awareness of tracking triggers privacy concerns that reduce effectiveness. This creates arms race between increasingly sophisticated tracking that remains invisible and consumer awareness that triggers resistance. Marketing manipulation increasingly relies on "dark patterns"—interface designs that trick users into unintended behaviors while maintaining illusion of control.

\subsection{Social Media as Manipulation Amplifier}

Social media platforms have fundamentally altered marketing manipulation dynamics through network effects, viral transmission, and parasocial relationships. Boyd and Ellison (2007) define social network sites as enabling users to construct public profiles, articulate connections, and traverse connection lists. While framed neutrally, these affordances create unprecedented manipulation opportunities through social proof amplification, influencer marketing, and viral fear propagation.

Marwick (2015) examines how Instagram creates "context collapse" where multiple audiences merge, intensifying self-presentation pressures. Her ethnographic research reveals how users internalize platform metrics—likes, follows, comments—as measures of self-worth. Brands exploit this metric fixation through engagement manipulation: buying followers, using engagement pods, and deploying bots to create false social proof. The 127 instances of social proof manipulation in our corpus likely underestimate total manipulation given these hidden amplification mechanisms.

Influencer marketing literature reveals sophisticated parasocial manipulation. Abidin (2016) traces the evolution from "micro-celebrity" to "influencer" as commercialization intensified. Her analysis shows how influencers perform authenticity through strategic self-disclosure, creating intimate connections that facilitate product promotion. The boundary between genuine recommendation and paid promotion deliberately blurs—our analysis found numerous instances where commercial relationships remained undisclosed despite legal requirements.

\subsection{Algorithmic Recommendation and Choice Architecture}

The role of algorithms in shaping consumer choice has attracted interdisciplinary attention spanning computer science, psychology, and critical theory. Seaver (2019) argues that recommendation algorithms don't merely predict preferences but actively construct them through feedback loops. His ethnographic study of music recommendation reveals how algorithms create "preference profiles" that become self-fulfilling prophecies—recommendations shape listening, which trains algorithms, which shape future recommendations.

Yeung (2017) introduces the concept of "hypernudge"—algorithmic nudging that operates continuously, dynamically, and pervasively. Unlike traditional nudges that preserve freedom of choice, hypernudges can effectively eliminate alternatives through extreme personalization. Her legal analysis warns that algorithmic manipulation may violate consumer protection laws, though enforcement remains minimal. The dynamic pricing and personalized offers we observe exemplify hypernudge techniques.

Bucher (2018) examines the "algorithmic imaginary"—how people understand and relate to algorithms despite their opacity. Her research reveals that consumers simultaneously overestimate algorithmic intelligence (attributing intention to statistical processes) and underestimate algorithmic influence (believing they maintain autonomous choice). This misunderstanding facilitates manipulation: consumers trust algorithmic recommendations as objective while remaining unaware of commercial optimization.

\section{Research Gaps and Thesis Contributions}
\label{sec:gaps}

\subsection{Integrative Analysis Across Sectors}

While substantial literature addresses marketing manipulation within specific sectors, comparative analysis across sectors remains limited. Fashion studies focus on luxury and identity, fitness research emphasizes body image and health, skincare literature examines beauty standards and aging—but systematic comparison revealing universal versus sector-specific patterns is absent. This thesis addresses this gap through parallel analysis of three sectors using identical analytical frameworks, enabling identification of both shared manipulation strategies and sector-specific variations.

The discovery that fear appeals dominate across all sectors (248 instances) despite different brand positioning reveals universal manipulation patterns obscured by sector-specific research. The fashion paradox—luxury brands employing fear more than aspiration—would remain hidden without cross-sector comparison. Similarly, the proliferation of scientific mimicry from skincare into fashion and fitness suggests strategy migration that sector-specific studies miss.

\subsection{Multimodal Manipulation in Digital Contexts}

Existing literature tends to analyze either textual or visual manipulation, rarely examining their interaction in digital environments. Text-focused studies miss visual manipulation, while visual analyses ignore linguistic strategies. Digital marketing operates through complex multimodal orchestration where meaning emerges from interaction between modes. This thesis addresses this gap through integrated multimodal analysis revealing how text, image, and interactive elements work synergistically to manipulate.

Our finding that brands systematically employ modal contradiction—aspirational imagery with fear-based text—demonstrates manipulation strategies invisible to single-mode analysis. The interactive dimension of digital marketing—quizzes, customization tools, social features—remains particularly understudied despite creating new manipulation possibilities through behavioral commitment and data extraction.

\subsection{Empirical Analysis of Manipulation Strategies}

While theoretical frameworks for understanding manipulation exist, empirical application to large-scale marketing corpora remains limited. Most studies analyze small samples or focus on single campaigns, preventing systematic understanding of manipulation patterns. This thesis analyzes 4.5 million characters of marketing text across 35 brands, providing unprecedented empirical foundation for understanding contemporary manipulation strategies.

The quantification of manipulation strategies—1,364 total instances categorized across 10 manipulation types and 8 emotional triggers—enables pattern identification impossible through qualitative analysis alone. The combination of quantitative pattern detection and qualitative discourse analysis reveals both scope and mechanisms of manipulation.

\subsection{Digital Amplification Mechanisms}

While digital transformation of marketing is widely acknowledged, specific mechanisms through which digital technologies amplify manipulation remain underexplored. Existing literature addresses components—personalization, tracking, algorithms—without systematic analysis of how they combine to intensify manipulation. This thesis identifies three amplification mechanisms: precision targeting, continuous optimization, and scale effects.

Our analysis reveals how these mechanisms interact: precision targeting identifies vulnerable moments, continuous optimization refines manipulation strategies, and scale effects enable simultaneous deployment across millions. This systematic understanding of digital amplification provides foundation for regulatory intervention and consumer protection.

\section{Methodological Approaches in Marketing Manipulation Research}
\label{sec:methods_lit}

\subsection{Quantitative Approaches: Experiments and Surveys}

Experimental methods dominate psychological research on marketing influence. Laboratory experiments provide causal evidence but suffer from ecological validity limitations—artificial settings may not reflect real-world marketing exposure. Field experiments offer greater realism but raise ethical concerns about manipulating consumers without consent. Our approach analyzes naturally occurring marketing texts, avoiding both validity and ethical limitations.

Survey research captures consumer perceptions and self-reported responses but faces social desirability bias—consumers may underreport susceptibility to manipulation. Response biases particularly affect sensitive topics like body image and financial pressure. Our text-based analysis examines what brands actually do rather than what consumers report experiencing, providing complementary perspective to survey research.

\subsection{Qualitative Approaches: Ethnography and Interview Studies}

Ethnographic research provides rich understanding of marketing consumption in cultural context. However, researcher presence may alter behavior, and findings from specific contexts may not generalize. Interview studies access consumer experiences and interpretations but rely on retrospective accounts that may be reconstructed or rationalized. Our discourse analysis examines marketing texts directly, capturing manipulation strategies independent of consumer interpretation.

\subsection{Mixed Methods: Combining Perspectives}

Mixed methods approaches combine quantitative and qualitative insights, addressing limitations of single methods. This thesis employs mixed methods through quantitative pattern detection (counting manipulation instances) and qualitative discourse analysis (interpreting meaning construction). The combination enables both systematic pattern identification and deep understanding of manipulation mechanisms.

The integration occurs at multiple levels: quantitative analysis identifies patterns warranting qualitative investigation, while qualitative insights inform quantitative categorization. This iterative process produces findings exceeding what either approach could achieve independently.

\section{Chapter Conclusion: Foundations for Critical Investigation}
\label{sec:lit_conclusion}

This literature review has mapped the scholarly landscape surrounding psychological manipulation in marketing discourse, revealing both substantial existing knowledge and critical gaps requiring investigation. The review demonstrates that while components of marketing manipulation have received attention—persuasion psychology, digital tracking, sector-specific studies—integrated understanding remains elusive.

Three major gaps emerge from this review. First, comparative analysis across sectors that reveals universal versus specific manipulation patterns remains absent. Second, multimodal analysis addressing the complex orchestration of textual, visual, and interactive elements in digital marketing is lacking. Third, large-scale empirical analysis documenting the scope and patterns of manipulation strategies has not been conducted.

This thesis addresses these gaps through comprehensive analysis of marketing discourse across fashion, fitness, and skincare sectors. By applying integrated theoretical framework combining Critical Discourse Analysis, psychological manipulation theory, and multimodal analysis to substantial empirical corpus, this research advances understanding of how contemporary marketing systematically exploits consumer vulnerabilities.

The literature also reveals methodological considerations shaping this investigation. The limitations of experimental and survey approaches—artificiality, ethical concerns, response biases—support our choice of naturalistic discourse analysis. The advantages of mixed methods—combining quantitative pattern detection with qualitative interpretation—inform our analytical approach.

Most significantly, the literature review establishes the urgent need for this research. As digital technologies amplify manipulation capabilities through micro-targeting, algorithmic optimization, and scale effects, understanding and addressing these practices becomes critical for consumer protection. The normalized exploitation revealed across sectors suggests systemic problem requiring comprehensive response.

The following chapter details the methodological framework developed to investigate these issues, building on insights from existing literature while addressing identified gaps. By establishing rigorous analytical procedures for examining marketing manipulation, this research provides empirical foundation for understanding and ultimately addressing the systematic exploitation of consumer vulnerabilities in digital marketing discourse.